---
title: "Week 2 exercises knitr and markdown"
author: "Vinay Mahajan"
date: "November 2, 2016"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### 1. Read the data file, set up the working directory and necessary libraries
```{r}

library(data.table)
library(ggplot2)
library(xtable)
setwd ("C:\\Users\\VinayMahajan\\Desktop\\Misc\\Coursera Data Science\\week2knitr\\source\\")

# 1 Create activity data
activity <- fread("activity.csv", check.names = TRUE)

activity <- activity[, visdate := as.POSIXct( gsub("-", "/", date), format="%Y/%m/%d") ]
activity <- activity[, weekdayend := ifelse(  format(visdate, "%u") %in% c(6, 7), "Weekend", "Weekday" )]

# Count number of records per day
activity0 <- activity [, nrow := .N , by = date]
activity0 <- activity0 [, nonmiss := sum(is.na(steps)) , by = date]

# Remove records where the non missing records are equal to all records
activity01 <- activity0 [ nrow > nonmiss, ]

### All the analysis will be carried out with days on which there is
### at least 1 non missing value for steps.

# 2 Histogram of the total number of steps taken each day
# Create Total steps for every day

activity2 <- activity01 [, .(sum = sum(steps),
                           mean = mean(steps),
                           median = median(steps) ) , by = visdate]

# Create the histogram with missing values
activity2_hist <- activity [, .(sum = sum(steps)) , by = visdate]
activity2_hist[is.na(activity2_hist)] <- 0
```

### 2. Histogram for total number of steps every day
```{r}
ggplot(activity2_hist, aes(x=sum)) + geom_histogram(binwidth=5000) +
         labs(title="Histogram of total number of steps every day",
              x="Total number of steps")

```

### 3. Mean and median number of days every day
```{r, message=FALSE}
print (activity2, type="html")
```

### 4. Time series plot of the average number of steps taken
```{r, message=FALSE}
ggplot( data = activity2, aes( visdate, mean)) + geom_line() +
         labs(title="Time series plot",
              x = "Dates",
              y = "Average number of steps taken")
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r, message=FALSE}
# The 5-minute interval that, on average, contains the maximum number of steps
activity3 <- activity01 [, .(mean = mean(steps)), by = interval]
activity3 <- activity3[order(-mean, interval)]
print (activity3, type="html")
```

### 6. Code to describe and show a strategy for imputing missing data
```{r, message=FALSE}
# Code to describe and show a strategy for imputing missing data
# Impute based on the average value for each time interval

activity4 <- activity [, mean := mean(steps,na.rm = TRUE), by = interval]
activity4 <- activity4 [, steps2 := ifelse(is.na(steps), mean, steps), ]


activity4_imp <- activity4 [, .(sum_imp = sum(steps2),
                             mean_imp = mean(steps2),
                             median_imp = median(steps2) ) , by = visdate]
```

### 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r, message=FALSE}
# Histogram based on imputed data
ggplot(activity4_imp, aes(x=sum_imp)) + geom_histogram(binwidth=5000) +
         labs(title="Histogram of total number of steps every day -- imputed",
              x="Total number of steps -- imputed")

```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r, message=FALSE}
# Panel plot comparing the average number of steps taken per
# 5-minute interval across weekdays and weekends
activity5 <- activity01[ , .(Avg_5int_Weekdays = mean(steps)), by= .(interval, weekdayend )]

qplot(interval, Avg_5int_Weekdays, 
      data = activity5, geom="line", facets = . ~ weekdayend) +
    labs(title="Panel plot: average number of steps taken weekday vs. weekend",
         x = "Interval", 
         y = "Average number of steps")
  
```
